---
title: "AdvancedRlab4: Linear Regression"
author: "Mariam Yayloyan"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AdvancedRlab4: Linear Regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The `AdvancedRlab4` package provides functions for fitting multiple linear
regression models. The main function, `linreg()`, fits a linear regression
model using the ordinary least squares (OLS) method. The package also
provides `linreg_qr()`, which fits a model using QR decomposition.

Both functions return objects of class `linreg`. Several S3 methods are
implemented for these objects, including `print()`, `summary()`, `coef()`,
`resid()`, `pred()`, and `plot()`.

This vignette demonstrates how to use the package with the built-in
`iris` dataset.

# Fitting a linear regression model

First, load the package:

```{r load-package}
library(AdvancedRlab4)
```

We use the `iris` dataset and model `Petal.Length` as a function of
`Sepal.Width` and `Sepal.Length`.

```{r fit-model}
model <- linreg(
  Petal.Length ~ Sepal.Width + Sepal.Length,
  data = iris
)
```

The `linreg()` function takes a formula describing the regression model
and a data frame containing the variables in the formula.
The returned object contains the estimated coefficients, fitted values,
residuals, residual variance, variance-covariance matrix, standard errors,
t-values, p-values, and other model information.

## Printing the model

The `print()` method provides a short overview of the fitted model.

```{r print-model}
print(model)
```

## Model summary

The `summary()` method provides more detailed information about the
fitted regression model.

```{r summary-model}
summary(model)
```

The coefficient table contains the estimated coefficients, standard
errors, t-values, and p-values. The summary also reports the residual
standard error and residual degrees of freedom.

## Extracting coefficients

The `coef()` method extracts the estimated regression coefficients.

```{r coef-model}
coef(model)
```

## Extracting residuals

The `resid()` method extracts the residuals from the fitted model.

```{r resid-model}
resid(model)
```

A residual is the difference between an observed response and its
corresponding fitted value.

## Obtaining fitted values

The `pred()` method returns the fitted values produced by the model.

```{r pred-model}
pred(model)
```

These values represent the predicted response for each observation in
the data used to fit the model.

## Diagnostic plots

The `plot()` method produces two diagnostic plots.

```{r plot-model}
plots <- plot(model)
```

The first plot shows residuals against fitted values.

```{r residual-plot, fig.width = 6, fig.height = 4}
plots$residual_plot
```

The second plot is a Scale-Location plot.

```{r scale-location-plot, fig.width = 6, fig.height = 4}
plots$scale_location_plot
```

These plots can be used to visually examine the residuals and identify
possible patterns in the model errors.

# QR decomposition implementation

The package also provides `linreg_qr()`, which fits a multiple linear
regression model using QR decomposition.

```{r fit-qr}
model_qr <- linreg_qr(
  Petal.Length ~ Sepal.Width + Sepal.Length,
  data = iris
)
```

The QR-based model is also an object of class `linreg`, so the same
methods can be used.

```{r summary-qr}
summary(model_qr)
```

The coefficients from the two implementations can be compared:

```{r compare-coef}
coef(model)
coef(model_qr)
```

The two implementations produce the same regression estimates up to
numerical precision, while using different computational approaches
to obtain the least-squares estimates.

# Conclusion

The `AdvancedRlab4` package provides an implementation of multiple
linear regression together with methods for inspecting and working
with fitted models. The `linreg()` function uses ordinary least
squares, while `linreg_qr()` uses QR decomposition.

Both functions return `linreg` objects that can be used with the
implemented `print()`, `summary()`, `coef()`, `resid()`, `pred()`, and
`plot()` methods.
